---
title: "Vaccination rates"
subtitle: "Data exploration"
author: "Chris Knox"
date: "`r Sys.Date()`"
output: 
  tint::tintHtml:
    toc: true
    toc_depth: 2
    number_sections: true
---

```{r setup, include=FALSE}
library(tint)
library(tidyverse)
library(ggthemes)
library(knitr)
library(scales)
library(RPostgreSQL)
library(r2d3)
library(ggrepel)
library(sf)

drv <- dbDriver('PostgreSQL')
con <- dbConnect(drv, dbname='vaccinations')

# invalidate cache when the package version changes
knitr::opts_chunk$set(tidy = FALSE,
                      cache.extra = packageVersion('tint'),
                      cache = TRUE,
                      autodep = TRUE,
                      echo = FALSE,
                      message = FALSE,
                      error = FALSE,
                      warning = FALSE
)

options(htmltools.dir.version = FALSE)
drake::r_make()
vaccinations <- drake::readd(oia_data)
vaccinations_geo <- drake::readd(oia_data_geo)
dom <- drake::readd(whole_dom)
```

# Introduction

This is the working document for a New Zealand Herald investigation of vaccination rates in New
Zealand - the final article is [here](https://www.nzherald.co.nz/nz/news/article.cfm?c_id=1&objectid=12250703)

This document has been released largely as-is and hopefully will provide some insight into and
transparency for the numerical aspects of the investigation. If you have any questions or comments
please get in touch with Chris Knox - https://twitter.com/vizowl or chris.knox@nzherald.co.nz

The actual document was created iteratively via duscussion between Kirsty Johnston
(https://twitter.com/kirsty_johnston) and Chris Knox as we tried to understand the data.

The source code for the analysis and interactive is open source. Although, if you use
it we would appreciate it if you could acknowledge the New Zealand Herald.

## Data

Immunisation data came from the Ministry of Health's [reporting](https://www.health.govt.nz/our-work/preventative-health-wellness/immunisation/immunisation-coverage/national-and-dhb-immunisation-data)
and from data released to the New Zealand Herald under the Official Information Act (OIA). The OIA'd
data is a breakdown of vaccination rates by Health Domicile - Health Domiciles are loosely similar
to suburbs.

This analysis is used data from StatsNZ - specifically area unit boundaries and geographic area
concordances.

All the government data is crown copyright released under a [creative commons licence](https://creativecommons.org/licenses/by/4.0/).


## Interactive

The charts were made as a simple [react](https://reactjs.org/) using the excellent 
[semiotic](https://semiotic.nteract.io/) visualisation library.

# Data from immunisation coverage reports

```{r,fig.fullwidth=T,fig.width=10,fig.height=14}
reports <- drake::readd(reports)
ggplot(reports %>% 
       filter(Measure=='Percentage',
              Breakdown=='Ethnicity',
              Category != 'Dep Unknown',
              DHB=='National'), aes(x=Quarter,y=Value,color=Category)) + 
  geom_line() + 
  facet_wrap(. ~ Milestone) +
  scale_color_brewer(palette='RdYlGn', direction=-1, name="Deprivation") +
  theme_fivethirtyeight() +
  labs(title="National vaccination milestones by ethnicity") 
```

```{r,fig.fullwidth=T,fig.width=10,fig.height=14}
ggplot(reports %>% 
       filter(Measure=='Percentage',
              Breakdown=='Deprivation',
              Category != 'Dep Unknown',
              DHB=='National'), aes(x=Quarter,y=Value,color=Category)) + 
  geom_line() + 
  facet_wrap(. ~ Milestone) +
  scale_color_brewer(palette='RdYlGn', direction=-1, name="Deprivation") +
  theme_fivethirtyeight() +
  labs(title="National vaccination milestones by deprivation") 
```


## Explore


```{r}
m6 <- drake::readd(month6)
m12 <- drake::readd(month12)
m18 <- drake::readd(month18)
m24 <- drake::readd(month24)
m60 <- drake::readd(month60)
worst6 <- top_n(m6, -1, Rate)
overall <- drake::readd(overall18) %>% ungroup
worstO <- top_n(overall, -1, Rate)
```

The worst vaccination rate at 6 months is `r percent(worst6 %>% pull(Rate))` in 
`r worst6 %>% pull(DOMICILE_NAME)` which is in `r worst6 %>% pull(DHB_NAME)`.


The worst overall (which might not mean anything) vaccination rate is `r percent(worstO %>% pull(Rate))` in 
`r worstO %>% pull(DOMICILE_NAME)` which is in `r worstO %>% pull(DHB_NAME)`.

Well that isn't so surprising - let's look at the worst 10.

```{r}
kable(top_n(m6, -10, Rate) %>% arrange(Rate) %>%
      select(DHB_NAME, DOMICILE_NAME, NUMBER_ELIGIBLE, Dep=CAU_average_NZDep2013, Rate),
    caption="The worst 10 domiciles at 6 months in 2018")
```

and overall.

```{r}
kable(top_n(overall, -10, Rate) %>% arrange(Rate) %>%
      select(DHB_NAME, DOMICILE_NAME, Dep=CAU_average_NZDep2013, Rate),
    caption="The worst 10 domiciles overall in 2018")
```

Let's use the score not the deciles.

```{r}
ggplot(m6, aes(x=CAU_average_NZDep_score_2013,y=Rate)) +
  geom_point() +
  geom_smooth() +
  theme_fivethirtyeight() +
  labs(title = "Deprivation versus 6 month vaccination rates") +
  scale_y_continuous(labels=percent)
```

How does it look by DHB

```{r,fig.fullwidth=T,fig.width=10,fig.height=14}
ggplot(m6, aes(x=CAU_average_NZDep_score_2013,y=Rate)) +
  geom_point() +
  geom_smooth() +
  theme_fivethirtyeight() +
  labs(title = "Deprivation versus 6 month vaccination rates") +
  scale_y_continuous(labels=percent,limits=c(0.5,1)) +
  facet_wrap(. ~ DHB_NAME, ncol=4)
```

Most of Auckland is better than most of Bay of Plenty.

How does it look at 12 months?


```{r,fig.fullwidth=T,fig.width=10,fig.height=14}
ggplot(m12, aes(x=CAU_average_NZDep_score_2013,y=Rate)) +
  geom_point() +
  geom_smooth() +
  theme_fivethirtyeight() +
  labs(title = "Deprivation versus 12 month vaccination rates") +
  scale_y_continuous(labels=percent,limits=c(0.5,1)) +
  facet_wrap(. ~ DHB_NAME, ncol=4)
```

```{r,fig.fullwidth=T,fig.width=10,fig.height=14}
ggplot(m18, aes(x=CAU_average_NZDep_score_2013,y=Rate)) +
  geom_point() +
  geom_smooth() +
  theme_fivethirtyeight() +
  labs(title = "Deprivation versus 18 month vaccination rates") +
  scale_y_continuous(labels=percent,limits=c(0.5,1)) +
  facet_wrap(. ~ DHB_NAME, ncol=4)
```

```{r,fig.fullwidth=T,fig.width=10,fig.height=14}
ggplot(m24, aes(x=CAU_average_NZDep_score_2013,y=Rate)) +
  geom_point() +
  geom_smooth() +
  theme_fivethirtyeight() +
  labs(title = "Deprivation versus 24 month vaccination rates") +
  scale_y_continuous(labels=percent,limits=c(0.5,1)) +
  facet_wrap(. ~ DHB_NAME, ncol=4)
```

```{r,fig.fullwidth=T,fig.width=10,fig.height=14}
ggplot(m60, aes(x=CAU_average_NZDep_score_2013,y=Rate)) +
  geom_point() +
  geom_smooth() +
  theme_fivethirtyeight() +
  labs(title = "Deprivation versus 60 month vaccination rates") +
  scale_y_continuous(labels=percent,limits=c(0.5,1)) +
  facet_wrap(. ~ DHB_NAME, ncol=4)
```

## Trends by deprivation

```{r}
ggplot(vaccinations %>%
       filter(AGE_IN_MONTHS == 6) %>%
       group_by(CAU_average_NZDep2013, MILESTONE_YEAR) %>%
       summarise(Rate=sum(NUMBER_FULLY_IMMUNISED)/sum(NUMBER_ELIGIBLE)),
     aes(x=MILESTONE_YEAR,y=Rate,color=coalesce(str_pad(CAU_average_NZDep2013, 2, pad='0'),'??'))) +
  geom_line() +
  scale_color_brewer(palette='RdYlGn', direction=-1, name="Deprivation") +
  theme_fivethirtyeight() +
  labs(title="6 month vaccination rates by deprivation") 
```


```{r}
ggplot(vaccinations %>%
       filter(AGE_IN_MONTHS == 12) %>%
       group_by(CAU_average_NZDep2013, MILESTONE_YEAR) %>%
       summarise(Rate=sum(NUMBER_FULLY_IMMUNISED)/sum(NUMBER_ELIGIBLE)),
     aes(x=MILESTONE_YEAR,y=Rate,color=coalesce(str_pad(CAU_average_NZDep2013, 2, pad='0'),'??'))) +
  geom_line() +
  scale_color_brewer(palette='RdYlGn', direction=-1, name="Deprivation") +
  theme_fivethirtyeight() +
  labs(title="12 month vaccination rates by deprivation") 
```

So it looks like drop-off is definitely occurring in more deprived areas. And 


# Load the data

## The data

The data covered the number of children who are vaccinated to each of the vaccination milestone by
health domicle and year.

It was provided by the Ministry of Health under the Official Information Act, as a spreadsheet with
68,385 rows and 8 columns. The columns are:

- **DHB_CODE**
- **DHB_NAME**
- **DOMICILE_CODE**
- **DOMICILE_NAME**
- **AGE_IN_MONTHS** - the age for each milestone
- **MILESTONE_YEAR** - the year the milestone
- **NUMBER_ELIGIBLE** - the number of children within the milestone age range
- **NUMBER_FULLY_IMMUNISED**

If there are less than ten eligible children in a domicile in a given year they have been aggregated
with other small domiciles in the DHB.


I haved used Stats NZ geographic areas data to match domicile codes
to 2013 area units and then too Otago's deprivation index data.

### Initial look

Calculate an overall rate for every domicile - this probably isn't something we
would use later - it is the average rate across all milestone ages and years.


```{r,fig.fullwidth=T,fig.width=10}
drake::readd(initialPlot1)
```
The scatter to the left and right around the individual deciles does not mean anything.
It is just _jitter_ that is randomly introduced to make it easier to read.

```{r,fig.fullwidth=T,fig.width=10,fig.height=20}
drake::readd(initialPlot2)
```

```{r,fig.height=20,fig.fullwidth=T,fig.width=10}
drake::readd(fitted_facets)
```

How does ethnicity effect things?

```{r}
ethnicity <- tbl(con, "census") %>% collect() %>%
  left_join(dom, by=c("code"="AU2013_code"))
ggplot(ethnicity, aes(x=Rate,y=maori,color=CAU_average_NZDep2013)) +
  geom_point()
```

Graphically I would have to say it doesn't ...



# For Emma

## National chart 

```{r}
ggplot(reports %>% filter(DHB=='National',Measure=="Percentage",Category=="Total"),
       aes(x=Quarter,y=Value, color=Milestone)) +
    geom_line() +
  scale_color_brewer(palette='Dark2', direction=-1) +
  theme_fivethirtyeight() +
  labs(title="New Zealand Immunisation Rates", subtitle="By Milestone Age") +
  scale_y_continuous(labels=percent) +
ggsave('nz-immunisation.png')
```

# More questions

## Dropoff by DHB

### 6 Months
```{r}
dhb_total <- reports %>%
  filter(Measure=='Percentage',Category=="Total",Breakdown=='Ethnicity') 
dhbNow <- dhb_total %>%
  filter(Quarter=='2019-03-01') %>% 
  select(DHB,Milestone,Now=Value)
dropoffs <- dhb_total %>%
  group_by(DHB, Milestone) %>%
  summarise(Max=max(Value,na.rm=T)) %>%
  ungroup() %>%
  inner_join(dhb_total %>% select(DHB, Milestone, Quarter, Value),
             by=c("DHB","Milestone","Max"="Value")) %>% 
  group_by(DHB,Milestone,Max) %>%
  summarise(`Best Quarter`=max(Quarter)) %>%
  inner_join(dhbNow, by=c("DHB","Milestone")) %>%
  mutate(Drop=Now-Max) %>%
  ungroup()

kable(dropoffs %>% filter(Milestone == '06 months') %>%
      arrange(Drop) %>%
      select(-Milestone) %>%
      mutate(Drop=percent(Drop),
             Max=percent(Max),
             Now=percent(Now)),
caption="Drop from each DHB's best 6 month immunisation rate until now")
```

### 8 Months
```{r}
kable(dropoffs %>% filter(Milestone == '08 months') %>%
      arrange(Drop) %>%
      select(-Milestone) %>%
      mutate(Drop=percent(Drop),
             Max=percent(Max),
             Now=percent(Now)),
caption="Drop from each DHB's best 8 month immunisation rate until now")

```




### 12 Months
```{r}
kable(dropoffs %>% filter(Milestone == '12 months') %>%
      arrange(Drop) %>%
      select(-Milestone) %>%
      mutate(Drop=percent(Drop),
             Max=percent(Max),
             Now=percent(Now)),
caption="Drop from each DHB's best 12 month immunisation rate until now")

```



## Dropoff by DHB and Ethnicity

### 6 Months
```{r}
dhb_total_eth <- reports %>%
  filter(Measure=='Percentage',Category != "Total",Breakdown=='Ethnicity') 
dhbNow_eth <- dhb_total_eth %>%
  filter(Quarter=='2019-03-01') %>% 
  select(DHB,Milestone,Now=Value,Category)
dropoffs_eth <- dhb_total_eth %>%
  group_by(DHB, Milestone,Category) %>%
  summarise(Max=max(Value,na.rm=T)) %>%
  ungroup() %>%
  inner_join(dhb_total_eth %>% select(DHB, Milestone, Quarter, Value,Category),
             by=c("DHB","Milestone","Max"="Value","Category")) %>% 
  group_by(DHB,Milestone,Max,Category) %>%
  summarise(`Best Quarter`=max(Quarter)) %>%
  inner_join(dhbNow_eth, by=c("DHB","Milestone","Category")) %>%
  mutate(Drop=Now-Max) %>%
  ungroup()

kable(spread(dropoffs_eth %>%
             filter(Milestone=='06 months') %>%
             select(DHB,Drop,Category), Category, Drop) %>%
      arrange(Māori) %>%
      mutate(Asian=percent(Asian),
             Māori=percent(Māori),
             Pākehā=percent(Pākehā),
             Other=percent(Other),
             Pacific=percent(Pacific)),
caption="Drop by ethnicity from each DHB's best 6 month immunisation rate until now")
```

### 8 Months

```{r}
kable(spread(dropoffs_eth %>%
             filter(Milestone=='08 months') %>%
             select(DHB,Drop,Category), Category, Drop) %>%
      arrange(Māori) %>%
      mutate(Asian=percent(Asian),
             Māori=percent(Māori),
             Pākehā=percent(Pākehā),
             Other=percent(Other),
             Pacific=percent(Pacific)),
caption="Drop by ethnicity from each DHB's best 8 month immunisation rate until now")
```



### 12 Months

```{r}
kable(spread(dropoffs_eth %>%
             filter(Milestone=='12 months') %>%
             select(DHB,Drop,Category), Category, Drop) %>%
      arrange(Māori) %>%
      mutate(Asian=percent(Asian),
             Māori=percent(Māori),
             Pākehā=percent(Pākehā),
             Other=percent(Other),
             Pacific=percent(Pacific)),
caption="Drop by ethnicity from each DHB's best 12 month immunisation rate until now")
```


## Dropoff by DHB and Deprivation

### 6 Months
```{r}
dhb_total_dep <- reports %>%
  filter(Measure=='Percentage',Breakdown=='Deprivation') 
dhbNow_dep <- dhb_total_dep %>%
  filter(Quarter=='2019-03-01') %>% 
  select(DHB,Milestone,Now=Value,Category)
dropoffs_dep <- dhb_total_dep %>%
  group_by(DHB, Milestone,Category) %>%
  summarise(Max=max(Value,na.rm=T)) %>%
  ungroup() %>%
  inner_join(dhb_total_dep %>% select(DHB, Milestone, Quarter, Value,Category),
             by=c("DHB","Milestone","Max"="Value","Category")) %>% 
  group_by(DHB,Milestone,Max,Category) %>%
  summarise(`Best Quarter`=max(Quarter)) %>%
  inner_join(dhbNow_dep, by=c("DHB","Milestone","Category")) %>%
  mutate(Drop=Now-Max) %>%
  ungroup()

kable(spread(dropoffs_dep %>%
             filter(Milestone=='06 months') %>%
             select(DHB,Drop,Category), Category, Drop) %>%
            arrange(`Dep 9-10`) %>%
            mutate(`Dep 1-2`=percent(`Dep 1-2`),
                    `Dep 3-4`=percent(`Dep 3-4`),
                    `Dep 5-6`=percent(`Dep 5-6`),
                    `Dep 7-8`=percent(`Dep 7-8`),
                    `Dep 9-10`=percent(`Dep 9-10`),
                    `Dep Unknown`=percent(`Dep Unknown`)),
caption="Drop by deprivation from each DHB's best 6 month immunisation rate until now")
```

### 8 Months

```{r}
kable(spread(dropoffs_dep %>%
             filter(Milestone=='08 months') %>%
             select(DHB,Drop,Category), Category, Drop) %>%
            arrange(`Dep 9-10`) %>%
            mutate(`Dep 1-2`=percent(`Dep 1-2`),
                    `Dep 3-4`=percent(`Dep 3-4`),
                    `Dep 5-6`=percent(`Dep 5-6`),
                    `Dep 7-8`=percent(`Dep 7-8`),
                    `Dep 9-10`=percent(`Dep 9-10`),
                    `Dep Unknown`=percent(`Dep Unknown`)
                    ),
caption="Drop by deprivation from each DHB's best 8 month immunisation rate until now")
```



### 12 Months

```{r}
kable(spread(dropoffs_dep %>%
             filter(Milestone=='12 months') %>%
             select(DHB,Drop,Category), Category, Drop) %>%
            arrange(`Dep 9-10`) %>%
            mutate(`Dep 1-2`=percent(`Dep 1-2`),
                    `Dep 3-4`=percent(`Dep 3-4`),
                    `Dep 5-6`=percent(`Dep 5-6`),
                    `Dep 7-8`=percent(`Dep 7-8`),
                    `Dep 9-10`=percent(`Dep 9-10`),
                    `Dep Unknown`=percent(`Dep Unknown`)),
caption="Drop by deprivation from each DHB's best 12 month immunisation rate until now")
```

# Comparisons

## Domicile's reaching 95%

What percentage of domicile's reach 95% vaccination by 12 months

```{r}
target12 <- m12 %>% count(DHB_NAME) %>%
  left_join(m12 %>% filter(Rate >= 0.95) %>% count(DHB_NAME),
                    by="DHB_NAME") %>%
      mutate(Frac=n.y/n.x) %>%
  select(-n.x,-n.y) %>% 
  arrange(Frac)
kable(target12 %>% mutate(Frac=percent(Frac)), caption = "Percentage of Domiciles that reach 95% by 12 months")

```

What are the characteristics of those domiciles?

```{r}
characteristics <- m12 %>%
  group_by(DHB_NAME) %>%
  summarise(DHB.DepScore=mean(CAU_average_NZDep_score_2013, na.rm=T),
            DHB.Size=mean(NUMBER_ELIGIBLE), DHB.Children =sum(NUMBER_ELIGIBLE))

missed_target12 <- m12 %>%
  filter(Rate < 0.95) %>%
  group_by(DHB_NAME) %>%
  summarise(Under.DepScore=mean(CAU_average_NZDep_score_2013, na.rm=T),
            Under.Size=mean(NUMBER_ELIGIBLE),
            Under.Children=sum(NUMBER_ELIGIBLE))

target12char <- characteristics %>%
  inner_join(m12 %>%
  filter(Rate >= 0.95) %>%
  group_by(DHB_NAME) %>%
  summarise(DepScore=mean(CAU_average_NZDep_score_2013, na.rm=T),
            Size=mean(NUMBER_ELIGIBLE),
            Children=sum(NUMBER_ELIGIBLE)),
  by = "DHB_NAME")

kable(characteristics %>% left_join(target12char, by="DHB_NAME") %>% arrange(DepScore))
kable(target12char)
```

Let's plot % target reached vs difference in deprivation score for those that reach the target
and those that don't

```{r}
ggplot(missed_target12 %>% left_join(target12char, by="DHB_NAME") %>%
       left_join(target12, by="DHB_NAME"),
     aes(x=DepScore-Under.DepScore,y=Frac)) +
  geom_point() +
  geom_text(aes(label=DHB_NAME),vjust=1.2) +
  theme_fivethirtyeight() +
  scale_y_continuous(labels=percent) +
  labs(title="Percentage of domiciles reaching target",
       subtitle="Versus the difference in DepScore")
ggsave('clusters.png')

```

Do the same plot but versus percentage of children - not domiciles.



```{r}
ggplot(missed_target12 %>% left_join(target12char, by="DHB_NAME") %>%
       left_join(target12, by="DHB_NAME"),
     aes(x=DepScore-Under.DepScore,y=Children/DHB.Children,label=paste(DHB_NAME,comma(DHB.Children-Children)))) +
  geom_point(colour="red") +
  theme_fivethirtyeight() +
  scale_y_continuous(labels=percent) +
  labs(title="Deprivation versus Vaccination",
       subtitle="Percentage of children fully vaccinated at 12 months versus the difference 
       in deprivation for target reaching and not-reaching areas",
      caption="Labels include number of children living in domiciles that haven't reached the target.") +
  geom_text_repel()
ggsave('clusters2.png')
```

Let's look at the averages for the acheived and failed domiciles by DHB

```{r,fig.height=12}
af <- missed_target12 %>%
  left_join(target12char, by="DHB_NAME") %>%
  left_join(target12, by="DHB_NAME") %>% 
  mutate(Achieved=Children/DHB.Children,Failed=Under.Children/DHB.Children) %>%
  select(DHB_NAME,Achieved,DepScore,Under.DepScore) %>%
  gather(Deprivation, Value, -Achieved, -DHB_NAME) %>%
  mutate(Status=case_when(Deprivation=="DepScore" ~ "Achieved", TRUE ~ "Not Achieved"),
         Children=case_when(Deprivation=="DepScore" ~ Achieved, TRUE ~ 1-Achieved),
         Deprivation=Value) %>%
  select(-Value,-Achieved)

ggplot(af,
     aes(x=Deprivation,y=Children,label=DHB_NAME)) +
  geom_point(aes(color=Status)) +
  geom_line(aes(group=DHB_NAME)) + 
  theme_fivethirtyeight() +
  scale_y_continuous(labels=percent) +
  labs(title="Deprivation versus Vaccination",
       subtitle="Percentage of children fully vaccinated at 12 months versus the difference 
       in deprivation for target reaching and not-reaching areas",
      caption="Labels include number of children living in domiciles that haven't reached the target.") +
  facet_wrap(. ~ DHB_NAME)
ggsave('clusters3.png', height=12)
```

What do these look like at 6 months?


```{r}
target6 <- m6 %>% count(DHB_NAME) %>%
  left_join(m6 %>% filter(Rate >= 0.95) %>% count(DHB_NAME),
                    by="DHB_NAME") %>%
      mutate(Frac=n.y/n.x) %>%
  select(-n.x,-n.y) %>% 
  arrange(Frac)
characteristics6 <- m6 %>%
  group_by(DHB_NAME) %>%
  summarise(DHB.DepScore=mean(CAU_average_NZDep_score_2013, na.rm=T),
            DHB.Size=mean(NUMBER_ELIGIBLE), DHB.Children =sum(NUMBER_ELIGIBLE))

missed_target6 <- m6 %>%
  filter(Rate < 0.95) %>%
  group_by(DHB_NAME) %>%
  summarise(Under.DepScore=mean(CAU_average_NZDep_score_2013, na.rm=T),
            Under.Size=mean(NUMBER_ELIGIBLE),
            Under.Children=sum(NUMBER_ELIGIBLE))

target6char <- characteristics6 %>%
  left_join(m6 %>%
  filter(Rate >= 0.95) %>%
  group_by(DHB_NAME) %>%
  summarise(DepScore=mean(CAU_average_NZDep_score_2013, na.rm=T),
            Size=mean(NUMBER_ELIGIBLE),
            Children=sum(NUMBER_ELIGIBLE)),
  by = "DHB_NAME")

```


```{r}
ggplot(missed_target6 %>% left_join(target6char, by="DHB_NAME") %>%
       left_join(target6, by="DHB_NAME") %>%
       filter(DHB_NAME != "Overseas and undefined"),
     aes(x=coalesce(DepScore-Under.DepScore,0),y=coalesce(Children/DHB.Children,0),label=paste(DHB_NAME,comma(DHB.Children-coalesce(Children,0))))) +
  geom_point(colour="red") +
  theme_fivethirtyeight() +
  scale_y_continuous(labels=percent, limits=c(-0.02,0.1)) +
  labs(title="Deprivation versus Vaccination",
       subtitle="Percentage of children fully vaccinated at 6 months versus the difference 
       in deprivation for target reaching and not-reaching areas",
      caption="Labels include number of children living in domiciles that haven't reached the target.") +
  geom_text_repel()
ggsave('clusters4.png')
```

# Population denisty

```{r,fig.fullwidth=T,fig.width=10,fig.height=14}
ggplot(vaccinations_geo %>% filter(AGE_IN_MONTHS==6),
       aes(x=log(NUMBER_ELIGIBLE/AREA_SQ_KM),y=NUMBER_FULLY_IMMUNISED/NUMBER_ELIGIBLE,color=CAU_average_NZDep_score_2013)) + 
                                                    geom_point() + 
                                                    facet_wrap(. ~ DHB_NAME) + 
                                                    scale_color_viridis_c(option="C",direction=-1,position="bottom")
```

It does not look like population density (which I think is an OK proxy for remoteness is a factor).


```{r}
missing <- reports %>% filter(Milestone=='06 months',Category=='Total',DHB=='National',Measure!='Percentage') %>% spread(Measure,Value) %>% mutate(Missing=Eligible-Immunised)
small_areas6nz <- vaccinations %>% filter(str_detect(DOMICILE_NAME, 'other domiciles'), AGE_IN_MONTHS==6) %>% group_by(MILESTONE_YEAR) %>% summarise(Total=sum(NUMBER_ELIGIBLE), Missed=sum(NUMBER_ELIGIBLE)-sum(NUMBER_FULLY_IMMUNISED))
```

# Maps

```{r}
vaccinations_geo <- vaccinations_geo %>%
  filter(AGE_IN_MONTHS==6,MILESTONE_YEAR==2016) %>%
  mutate(OK=case_when(NUMBER_FULLY_IMMUNISED/NUMBER_ELIGIBLE >= 0.75 ~ 'Yes', TRUE ~ 'No'))
ggplot(vaccinations_geo %>% filter(DHB_NAME=='Bay of Plenty'),
       aes(geometry=geometry,fill=OK)) + 
  geom_sf() + 
  scale_fill_viridis_d(option='C',name="Rate") + 
  theme_map() +
  labs(subtitle="Bay of Plenty at 6 months", title="75% of children vaccinated")
```

```{r}
ggplot(vaccinations_geo %>% filter(DHB_NAME=='Counties Manukau'),
       aes(geometry=geometry,fill=OK)) + 
  geom_sf() + 
  scale_fill_viridis_d(option='C',name="Rate") + 
  theme_map() +
  labs(subtitle="Counties Manukau at 6 months", title="75% of children vaccinated")
```

```{r}
ggplot(vaccinations_geo %>% filter(DHB_NAME=='Waikato'),
       aes(geometry=geometry,fill=OK)) + 
  geom_sf() + 
  scale_fill_viridis_d(option='C',name="Rate") + 
  theme_map() +
  labs(subtitle="Waikato at 6 months", title="75% of children vaccinated")
```

```{r}
ggplot(vaccinations_geo %>% filter(DHB_NAME=='Northland'),
       aes(geometry=geometry,fill=OK)) + 
  geom_sf() + 
  scale_fill_viridis_d(option='C',name="Rate") + 
  theme_map() +
  labs(subtitle="Northland at 6 months", title="75% of children vaccinated")
```


# Numbers

## What is the difference between December 2015 and now?

670 kids - that is all that are missing.

Peak 6 month overall rate was 82% - now it is 77%. In the most recent quarter there were
14745 kids - so vaccinating 670 more would get to 82%.

## Demographics of drop.

Deprivation

```{r}
mdall6 <- reports %>%
     filter(Breakdown=='Deprivation', Milestone=='06 months',DHB=='National')
kable(mdall6 %>%
     filter(Measure=="Percentage",Quarter=="2015-12-01") %>%
     select(Category, Value) %>%
     inner_join(mdall6 %>%
     filter(Measure=="Percentage",Quarter=="2019-03-01") %>%
     select(Category, Value), by="Category") %>%
     mutate(Drop=Value.x-Value.y) %>%
     inner_join(mdall6 %>%
     filter(Measure=="Eligible",Quarter=="2015-12-01") %>%
     select(Category, Value), by="Category") %>%
     mutate(Lost=round(Value*Drop)) %>%
     mutate(Drop=round(Drop*100,2),`Dec 2015`=percent(Value.x),`Mar 2019`=percent(Value.y)) %>%
     select(-Value.x,-Value.y) %>%
     rename(Eligible=Value),
  caption="Additional number of children that would have been vaccinated at 6 months in March this
year if vaccination rate was the same as Dec 2015", align=c('l', rep('r',5)))
```

Ethnicity


```{r}
meall6 <- reports %>%
     filter(Breakdown=='Ethnicity', Milestone=='06 months',DHB=='National')
kable(meall6 %>%
     filter(Measure=="Percentage",Quarter=="2015-12-01") %>%
     select(Category, Value) %>%
     inner_join(meall6 %>%
     filter(Measure=="Percentage",Quarter=="2019-03-01") %>%
     select(Category, Value), by="Category") %>%
     mutate(Drop=Value.x-Value.y) %>%
     inner_join(meall6 %>%
     filter(Measure=="Eligible",Quarter=="2015-12-01") %>%
     select(Category, Value), by="Category") %>%
     mutate(Lost=round(Value*Drop)) %>%
     mutate(Drop=round(Drop*100,2),`Dec 2015`=percent(Value.x),`Mar 2019`=percent(Value.y)) %>%
     select(-Value.x,-Value.y) %>%
     rename(Eligible=Value),
  caption="Additional number of children that would have been vaccinated at 6 months in March this
year if vaccination rate was the same as Dec 2015", align=c('l', rep('r',5)))
```

# DHB Numbers

Deprivation

```{r}
mdall6 <- reports %>%
     filter(Breakdown=='Deprivation', Milestone=='06 months',DHB != 'National')
kable(mdall6 %>%
     filter(Measure=="Percentage",Quarter=="2015-12-01") %>%
     select(Category, Value, DHB) %>%
     inner_join(mdall6 %>%
     filter(Measure=="Percentage",Quarter=="2019-03-01") %>%
     select(Category, Value, DHB), by=c("Category","DHB")) %>%
     mutate(Drop=Value.x-Value.y) %>%
     inner_join(mdall6 %>%
     filter(Measure=="Eligible",Quarter=="2015-12-01") %>%
     select(Category, Value, DHB), by=c("Category","DHB")) %>%
     mutate(Lost=round(Value*Drop)) %>%
     mutate(Drop=round(Drop*100,2),`Dec 2015`=percent(Value.x),`Mar 2019`=percent(Value.y)) %>%
     select(-Value.x,-Value.y) %>%
     rename(Eligible=Value) %>%
     arrange(DHB, Category),
  caption="Additional number of children that would have been vaccinated at 6 months in March this
year if vaccination rate was the same as Dec 2015", align=c('l', rep('r',5)))
```

Ethnicity


```{r}
mdall6 <- reports %>%
     filter(Breakdown=='Ethnicity', Milestone=='06 months',DHB != 'National')
kable(mdall6 %>%
     filter(Measure=="Percentage",Quarter=="2015-12-01") %>%
     select(Category, Value, DHB) %>%
     inner_join(mdall6 %>%
     filter(Measure=="Percentage",Quarter=="2019-03-01") %>%
     select(Category, Value, DHB), by=c("Category","DHB")) %>%
     mutate(Drop=Value.x-Value.y) %>%
     inner_join(mdall6 %>%
     filter(Measure=="Eligible",Quarter=="2015-12-01") %>%
     select(Category, Value, DHB), by=c("Category","DHB")) %>%
     mutate(Lost=round(Value*Drop)) %>%
     mutate(Drop=round(Drop*100,2),`Dec 2015`=percent(Value.x),`Mar 2019`=percent(Value.y)) %>%
     select(-Value.x,-Value.y) %>%
     rename(Eligible=Value) %>%
     arrange(Category, DHB),
  caption="Additional number of children that would have been vaccinated at 6 months in March this
year if vaccination rate was the same as Dec 2015", align=c('l', rep('r',5)))
```


# Final charts

```{r}
ggplot(reports %>% 
       filter(Milestone=='06 months',
              Measure=='Percentage',
              Breakdown=='Deprivation',
              Category != 'Dep Unknown',
              DHB=='National'), aes(x=Quarter,y=Value,color=Category)) + 
  geom_line() + 
  facet_wrap(. ~ Milestone) +
  scale_y_continuous(labels=percent) +
  scale_color_brewer(palette='RdYlGn', direction=-1, name="Deprivation") +
  theme_fivethirtyeight() +
  labs(title="Fully vaccinated six month olds") 
ggsave('deprivation.svg')
```

```{r}
ggplot(reports %>% 
       filter(Milestone=='06 months',
              Measure=='Percentage',
              Breakdown=='Ethnicity',
              DHB=='National'), aes(x=Quarter,y=Value,color=Category)) + 
  geom_line() + 
  facet_wrap(. ~ Milestone) +
  scale_y_continuous(labels=percent) +
  scale_color_brewer(palette='RdYlGn', direction=-1, name="Ethnicity") +
  theme_fivethirtyeight() +
  labs(title="Fully vaccinated six month olds") 
ggsave('ethnicity.svg')
```

```{r}
ggplot(reports %>% 
       filter(Milestone=='06 months',
              Measure=='Percentage',
              Breakdown=='Ethnicity',
              Category == 'Total',
              DHB=='National'), aes(x=Quarter,y=Value,color=Category)) + 
  geom_line() + 
  facet_wrap(. ~ Milestone) +
  scale_y_continuous(labels=percent) +
  scale_color_brewer(palette='RdYlGn', direction=-1) +
  theme_fivethirtyeight() +
  theme(legend.position="none") +
  labs(title="Fully vaccinated six month olds") 
ggsave('total.svg')
```

